Stateless Model Checking with Data-Race Preemption Points

### 2015-10-10 pldi-16/di.tex
tighten up prior commit and write a conclusion

@@

@@

When \landslide~detects a data race, it produces a report for each of the two memory accesses involved in the race.
Each report indicates the program counter value (PC) associated with the access, as well as some further conditions to help filter away unrelated executions of the same instruction on different data.
(For example, {\tt list\_insert()} might be called from many parts of a codebase, but produce data-race candidates from only one callsite.)
Ideally, the PC would be qualified by a full backtrace, but tracing the stack is too expensive to do as often as each shared memory access.
Instead, \landslide~qualifies the PC with
(a) the current thread ID and
(b) the most recent {\tt call} instruction.
% (a crude approximation of a stack trace)
% which are much cheaper, as we carry them around all the time already
Note that we do {\em not} qualify data races by {\em memory addresses},
which change across thread interleavings.
% especially when malloc is involved.
Figure~\ref{fig:dont-filter-dr-by-address} shows example code where qualifying by memory address will miss the bug.

% TODO: make dont-filter-dr-by-address example.

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
Rather than adding a maximal job with both new PPs at once, we prefer to add smaller jobs which have a higher chance of completing in time.
If those state spaces are bug-free, they will in turn add the maximal job later.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance: while the big job may not complete in time, the small job might miss the data race entirely if the original PPs were required to expose it.
In practice, we have observed some bugs found quickly by the small job, and other bugs missed by the small job found eventually by the big job.
% which justifies this design

>>>>

\subsection{Filtering false positive data races}

% TODO: proof goes here
% TODO: example codez with malloc re free go here

@@

\quicksand~uses a workqueue model
...
priorities... data race priorities...

@@

When \landslide~detects a data race, it reports each of the two memory accesses involved in the race.
Each report indicates the program counter value (PC) associated with the access, as well as some further conditions to help filter away unrelated executions of the same instruction on different data.
(For example, {\tt list\_insert()} might be called from many parts of a codebase, but produce data-race candidates from only one callsite.)
Ideally, the PC would be qualified by a full backtrace, but tracing the stack is too expensive to do for each shared memory access.
Instead, \landslide~qualifies the PC with
(a) the current thread ID and
(b) the most recent {\tt call} instruction.
% (a crude approximation of a stack trace)
% which are much cheaper, as we carry them around all the time already
Note that we do {\em not} qualify data races by the shared memory address,
which can change based on different interleavings of previous code
(for example, depending on the result of {\tt malloc()}).
% especially when malloc is involved.
% TODO: decide whether to make dont-filter-dr-by-address example.
%Figure~\ref{fig:dont-filter-dr-by-address} shows example code where qualifying by memory address will miss the bug.

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
Rather than adding a maximal job with both new PPs at once, we prefer to add smaller jobs which have a higher chance of completing in time.
If those state spaces are bug-free, they will in turn add the maximal job later.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance:
while the big job risks not completing in time,
the small job risks missing the data race entirely if the original PPs were required to expose it.
In practice, we observed some bugs found quickly by these small jobs, and other bugs missed by the small jobs found eventually by the big jobs.
% which justifies this design

### 2015-10-10 pldi-16/paper.tex
some touchups and talk about test suite
tighten up prior commit and write a conclusion

@@

% TODO: Publish source of ID wrapper alone, if can't publish landslide itself.
% TODO: Publish log files.

@@

Many thanks to David A. Eckhardt, Vaishaal Shankar, and Haryadi Gunawi for generously providing student implementations from CMU's, Berkeley's, and U. of Chicago's OS classes respectively.
Thanks to Ji\v{r}\'{i} \v{S}im\v{s}a and the anonymous PLDI reviewers for their helpful comments.
This work was supported in part by % TODO: funding sources

>>>>

% TODO: Finalize these numbers.
\newcommand\numthrlibs{{\Large \bf NUMP2S}}
\newcommand\numpintoses{{\Large \bf NUMPINTOSES}}
\newcommand\numstudence{{\Large \bf NUMSTUDENCE}} % total pintoses plus p2s

@@

We have presented Iterative Deepening, a new technique for automating the choice of preemption points (PPs) during stateless model checking, and \quicksand, a tool which implements this technique tailored to the \landslide~model checker.
\quicksand~incorporates \landslide's data-race analysis to create new PPs tailored specifically to the program under test,
and manages multiple \landslide~instances to test many different PP subsets in a given CPU budget, even when the full state space of all PPs would be computationally intractable.

Our evaluation shows that \quicksand~achieves better bug-finding results than either single-pass data-race detection or single-state-space model checking (SSS-MC) alone.
We avoid all false-positive data race reports by verifying the absence of bugs in their associated state spaces, and find crashes caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
Compared to SSS-MC, we find bugs faster in subset state spaces when the maximal state space is too large to explore, and find new bugs with data-race PPs that the maximal state space would not expose at all.

Although \landslide~itself is not publicly available due to its dependency on the non-free simulator {\sc Simics}, \quicksand~is open-source and its interface can be adapted to fit any similar tool.
We have also posted the log files from all our experiments.
They can be found here:

\url{https://github.com/bblum/landslide} % TODO: put QS in its own repository

\url{https://github.com/bblum/landslide} % TODO: put final logs somewhere

@@

Many thanks to David A. Eckhardt, Vaishaal Shankar, and Haryadi Gunawi for generously providing student implementations from CMU's, Berkeley's, and U. of Chicago's OS classes respectively.
Thanks to Wind River Simics for %TODO
Thanks to Ji\v{r}\'{i} \v{S}im\v{s}a and the anonymous PLDI reviewers for their helpful comments.
This work was supported in part by % TODO: funding sources
